Compositional equivalences based on Open pNets\thanks{This work was partially funded by the Associated Team FM4CPS between INRIA and ECNU, Shanghai}

### 2019-05-09 2016-Weak-FH-Bisim/WeakBisim.tex
skype call

@@

For the sake of simplicity, we will  restrict $n$ and $m$ to be equal to $1$, hence a single tau open transition will be considered; however the proof may be easily generalized to the multiple tau open transitions by using the same reasoning and (\WTTrois)~rule. Consider each open transition separately:
\begin{enumerate}
\item For the first open transition in $\mathcal{T}_1$:
\begin{mathpar}
%\inferrule{
 \openrule
    {
       \beta_{1j}^{j\in J_1'},\Pred_1,\Post_1   }
         {s_{10} \OTarrow {\tau} s_{11}}          
\end{mathpar}
we have $(s,t|\Pred_{s,t})\in\mathcal{R}$ by hypothesis, and $s=s_{10}$ the we have $(s_{10},t|\Pred_{s_{10},t})\in\mathcal{R}$. Thus, by  Definition \ref{def-Weak-bisim} we  can deduce there exist weak open transitions $\symb{WOT}_x^{x\in X} \subseteq \WT_2$:
 \begin{mathpar}
    \openrule
         {
           \gamma_{j x}^{j\in J_{x}}, \Pred_{OT_x},\Post_{OT_x}}
         {t \OTWeakarrow {\alpha_x} t_x}
\end{mathpar}
 such that  $\forall x,J_{x}=\{j\in J_1'|\beta_{1j}\neq\tau\}, (s_{11},t_x|\Pred_{s_{11},t_x})\in \mathcal{R}$. \\
 $\Pred_{s,t} \land \Pred_1\\
\hspace{1cm} \implies\!\!\! \displaystyle{\bigvee_{x\in X}}
  \!\! \left( \forall j\in J_x.\vis{\beta_{1j}}=\gamma_{jx}  \Rightarrow \Pred_{OT_x}
     \land \alpha_x\!=\!\tau \land  
     \Pred_{s_{11},t_x}\subst{\Post_1\uplus\Post_{OT_x}}\right)$

@@

\Big(\forall j\in J_x. \vis{\beta_{1j}}\!\!=\!\!\gamma_{jx} \land  \forall j \in J_{x y}. \vis{\beta_{2j}}\subst{\Post_1\uplus\Post_{OT_x}} \!=\!\gamma_{jx y}  \Rightarrow\Pred_{OT_x} \land  \Pred_{OT_{x y}} \subst{\Post_1\uplus\Post_{OT_x}} \land \alpha_x\!=\!\tau \land \alpha\!=\!\alpha_{x y} \subst{\Post_1\uplus\Post_{OT_x}} \land \Pred_{s_{21},t_{x y}}\subst{\Post_2\shortotimes\Post_1\uplus \Post_{OT_{x y}}\shortotimes \Post_{OT_x}}\Big)$\\ 

@@

\begin{lemma}[Weak OT decomposition]\label{lem-decomposeWOT} 
	Let $\Leaves(Q)=pLTS_l^{l\in L_Q}$; suppose\footnote{Note that the hypotheses of the 
	lemma imply that $Q$ is 
	not a \pLTS\ but a similar lemma can be proven for a pLTS $Q$}:
	\[ P[Q]_{j_0}  
		\models
		{\openrule
			{
				\gamma_j^{j\in J}, \Pred,  
				\Post}
			{\ostate{s_i^{i\in L}} \OTWeakarrow {\alpha}
				\ostate{s_i'^{\, i\in L}}}
		}
	\]
		with  $J\cap\Holes(Q)\neq\emptyset$ or $\exists i\in L_Q.\,s_i\neq s'_i$, i.e. $Q$ takes part in the reduction.  
		 Then there exist $\gamma_{j_0}=[\beta_0..\beta_n]$, $\Pred\,'$,  
		$\Post\,'$,   and for all $p\in[0..n]$ there exist $\alpha_p$, $\Pred_p$, $\Post_p$ and a family $\gamma_{p j}^{j\in J_p}$ and for all $p\in[0..n+1]$ $s_{p i}^{\,i\in L_Q}$. s.t.:\\[-2ex]
		%\[
		\begin{mathpar}
		P\models{\openrule
			{
				\gamma_j^{j\in (J_p\setminus \Holes(Q)) \cup \{j_0\}}, 
				\Pred\,',  
				\Post\,'}
			{\ostate{s_i^{i\in L\setminus L_Q}} \OTWeakarrow {\alpha}
				\ostate{s_i'^{\,i\in L\setminus L_Q}}}
		}%\]
	\vspace{-2.2ex}\\\text{and for all $p\in[0..n]$~~}
		%\[
		Q\models{\openrule
			{
				\gamma_{p j}^{j\in J_p}, \Pred_p,  
				\Post_p}
			{\ostate{s_{p i}^{i\in L_Q}} \OTWeakarrow {\alpha_p}
				\ostate{s_{(p+1) i}^{\,i\in L_Q}}}
		}%\]
\\
		\text{and~~}  \bigcup_{p=0}^nJ_{p} = J\cap\Holes(Q) \qquad
 \text{and\qquad} \forall j\in J_p,\,\gamma_j= \mybigdotcup_{p=0}^n \gamma_{p j} \\
 \text{and\quad} \forall i \in L_Q.\, s_{(n+1) i} = s_i'\land s_{0 i} = s_i\\{\hspace{-3ex}}\text{and~~} \Pred \iff \Pred\,'
		\land \!\!\bigwedge_{p=0}^n((\alpha_p= \beta_p\land \Pred_p)\bigotimes_{i=0}^{p-1}\Post_i) 
\\
\text{and~}
\Post=\Post\,' \uplus \bigotimes_{p=0}^n
		\Post_p  
		\end{mathpar}
where $\Post_p$ only act upon variables of 
		$\Leaves(Q)$.
\end{lemma}

@@

Consequently, for  $\gamma_{j_0}=[\beta_0..\beta_n]$ such that $\beta_i=\alpha'_i$ we have $n$ weak open transitions  $\symb{WOT}_k$ such that $k\in [0..n]$.
Therefore, we can conclude that we have for all $k\in [0..n]$:  
\begin{mathpar}
Q\models{\openrule
			{
			\gamma_{kj}^{j\in J_k}, {\Pred_k}, {\Post_k}}
			{\ostate{s_{pl}^{l\in L_Q}} \OTWeakarrow {\alpha_{k}}
				\ostate{s_{(p+1)l}^{l\in L_Q}}}
		}%\]
\end{mathpar}
and $\displaystyle{\bigcup_{p=0}^nJ_{p}} = J\cap\Holes(Q)$, $\forall j\in J_p,\,\gamma_j= \displaystyle{\mybigdotcup_{p=0}^n \gamma_{p j}}$, $\forall l\in L_Q,\, s_{0l}=s_l$, $s_{(n+1)l}=s'_l$  \\
  $\Pred \iff \Pred\,'
		\land \displaystyle{\bigwedge_{p=0}^n}((\alpha_p= \beta_{pj_0}\land \Pred_p)\bigotimes_{i=0}^{p-1}\Post_i),\text{and~}$\\
$\displaystyle{\Post=\Post\,'\uplus\bigotimes_{p=0}^n
		\Post_p}$  \\
We have:\\
$\Pred=\displaystyle{\bigwedge_{i=0}^{m_1}\Pred_{1i}\subst{\bigotimes_{j=i-1}^{0}\!\!\Post_{1j}}\land\Pred_2 \subst{\bigotimes_{j=m_1}^{0}\!\!\Post_{1j}}\land 
\bigwedge_{i=0}^{m_2}\Pred_{3i}\subst{\bigotimes_{j=i-1}^{0}\!\!\Post_{3j}\shortotimes\Post_2\shortotimes\bigotimes_{j=m_1}^{0}\!\!\Post_{1j}}}$\\
By replacing $\Pred_{1i}$,  $\Pred_2$, and $\Pred_{3i}$ by their equivalent formula  we obtain:\\
$\Pred \iff\displaystyle{\bigwedge_{i=0}^{m_1}({\Pred'_{1i}}
		\land {\Pred''_{1i}}\land \alpha_{1i}=\beta_{1ij_0}) \bigotimes_{j=0}^{i-1}\subst{\Post_{1j}}\land}\\
~~\qquad\qquad\qquad({\Pred'_2}\land {\Pred''_2}\land \alpha_2=\beta_{2j_0})	\bigotimes_{i=0}^{m_1}\subst{\Post_{1i}}\land\\
~~\qquad\qquad\bigwedge_{i=0}^{m_2} ({\Pred'_{3i}}\land {\Pred''_{3i}}\land \alpha_{3i}=\beta_{3ij_0})
\subst{\bigotimes_{j=i-1}^{0}\!\!\Post_{3j}\shortotimes\Post_2\shortotimes\bigotimes_{j=m_1}^{0}\!\!\Post_{1j}}$\\
By distribution of the substitution operation over the logical operator we can deduce:\\
$\Pred\iff \displaystyle{\big(\bigwedge_{i=0}^{m_1}
{\Pred'_{1i}}\bigotimes_{j=0}^{i-1}\subst{{\Post'_{1j}}}\big)
\land{\Pred'_2}\bigotimes_{i=0}^{m_1}\subst{{\Post'_{1i}}}\land} \\
~~\qquad\qquad\quad\big(\bigwedge_{i=0}^{m_2}{\Pred'_{3i}}\bigotimes_{j=0}^{i-1}\subst{{\Post'_{3j}}}\subst{{\Post'_2}}\bigotimes_{j=0}^{m_1}\subst{{\Post'_{1j}}}\big) \land\\	
~\qquad\qquad\quad\bigwedge_{i=0}^{m_1}({\Pred''_{1i}}\land \alpha_{1i}=\beta_{1ij_0}) \bigotimes_{j=0}^{i-1}\subst{\Post_{1j}}\land({\Pred''_2}\land \alpha_2=\beta_{2j_0})	\bigotimes_{i=0}^{m_1}\subst{\Post_{1i}}\land\\
~\qquad\qquad\quad\bigwedge_{i=0}^{m_2} ({\Pred''_{3i}}\land \alpha_{3i}=\beta_{3ij_0})\bigotimes_{j=0}^{i-1}\subst{\Post_{3j}}\subst{\Post_2}  \bigotimes_{i=0}^{m_1}\subst{\Post_{1i}}
$\\		
Therefore, we have:\\
$\Pred\iff \Pred\,' \wedge \displaystyle{\bigwedge_{p=0}^{n}}((\alpha_p= \beta_{pj_0}\land \Pred_p)\bigotimes_{i=0}^{p-1}\Post_i)$\\ where $n=m_1+m_2+1$	\\
Concerning the postcondition we have:\\
$\Post=\displaystyle{\bigotimes_{j=m_2}^{0}\!\!\Post_{3j}\shortotimes\Post_2\shortotimes\bigotimes_{j=m_1}^{0}\!\!\Post_{1j}}$\\
By replacing ${\Post_{1i}}$,  $\Post_2$, and ${\Post_{3i}}$ and because the ${\Post'_i}$ and ${\Post''_i}$ are independent (the ${\Post'_i}$ concern the leaves of $P$ while the ${\Post''_i}$ concern the leaves of $Q$), we obtain:\\
$\Post= \displaystyle{\bigotimes_{i=0}^{m_2}\subst{{\Post'_{3i}}\uplus 
		{\Post''_{3i}}} \subst{{\Post'_2}\uplus 
		{\Post''_2}} \bigotimes_{j=0}^{m_1}\subst{ {\Post'_{1i}}\uplus 
		{\Post''_{1i}}}}\\
~~\qquad= \bigotimes_{i=0}^{m_2}{\Post'_{3i}}\subst{{\Post'_2}}\bigotimes_{j=0}^{m_1}\subst{{\Post'_{1j}}} \uplus \bigotimes_{i=0}^{m_2}{\Post''_{3i}}\subst{{\Post''_2}}\bigotimes_{j=0}^{m_1}\subst{{\Post''_{1j}}} \\
~~\qquad= \Post' \uplus \bigotimes_{j=0}^{n} \Post_j $\\
 where $n=m_1+m_2+1$	
\end{proof}

>>>>

Another consequence of the semantics and logical interpretation of the formulas is that we make no distinction between the equality and the equivalence on boolean formulas, i.e. equivalence of two predicates $\Pred$ and $\Pred\,'$ can be denoted $\Pred=\Pred\,'$.

@@

For the sake of simplicity, we prove the rule in the  restricted case where $n$ and $m$ to be equal to $1$, hence a single tau open transition will be considered on each side of the potentially visible one. The proof may be easily generalized to the multiple tau open transitions by using the same reasoning and (\WTTrois)~rule. Consider each open transition separately:
\begin{enumerate}
\item For the first open transition in $\mathcal{T}_1$:
\begin{mathpar}
%\inferrule{
 \openrule
    {
       \beta_{1j}^{j\in J_1'},\Pred_1,\Post_1   }
         {s_{10} \OTarrow {\tau} s_{11}}          
\end{mathpar}
we have $(s,t|\Pred_{s,t})\in\mathcal{R}$ by hypothesis, and $s=s_{10}$ the we have $(s_{10},t|\Pred_{s_{10},t})\in\mathcal{R}$. Thus, by  Definition \ref{def-Weak-bisim} we  can deduce there exist weak open transitions $\symb{WOT}_x^{x\in X} \subseteq \WT_2$:
 \begin{mathpar}
    \openrule
         {
           \gamma_{j x}^{j\in J_{x}}, \Pred_{OT_x},\Post_{OT_x}}
         {t \OTWeakarrow {\alpha_x} t_x}
\end{mathpar}
 such that  $\forall x,J_{x}=\{j\in J_1'|\beta_{1j}\neq\tau\}, (s_{11},t_x|\Pred_{s_{11},t_x})\in \mathcal{R}$. \\
 $\Pred_{s,t} \land \Pred_1\\
\hspace{1cm} \implies\!\!\! \displaystyle{\bigvee_{x\in X}}
  \!\! \left( \forall j\in J_x.\vis{\beta_{1j}}=\gamma_{jx}  \Rightarrow \Pred_{OT_x}
     \land \alpha_x\!=\!\tau \land  
     \Pred_{s_{11},t_x}\subst{\Post_1\uplus\Post_{OT_x}}\right)$

@@

\Big(\forall j\in J_x. \vis{\beta_{1j}}\!\!=\!\!\gamma_{jx} \land  \forall j \in J_{x y}. \vis{\beta_{2j}}\subst{\Post_1\uplus\Post_{OT_x}} \!=\!\gamma_{jx y}  \Rightarrow\Pred_{OT_x} \land  \Pred_{OT_{x y}} \subst{\Post_1\uplus\Post_{OT_x}} \land \alpha_x\!=\!\tau \land (\alpha\!=\!\alpha_{x y}) \subst{\Post_1\uplus\Post_{OT_x}} \land \Pred_{s_{21},t_{x y}}\subst{\Post_2\shortotimes\Post_1\uplus \Post_{OT_{x y}}\shortotimes \Post_{OT_x}}\Big)$\\ 

@@

\begin{lemma}[Weak OT decomposition]\label{lem-decomposeWOT} 
	Let $\Leaves(Q)=pLTS_l^{l\in L_Q}$; suppose\footnote{Note that the hypotheses of the 
	lemma imply that $Q$ is 
	not a \pLTS\ but a similar lemma can be proven for a pLTS $Q$}:
	\[ P[Q]_{j_0}  
		\models
		{\openrule
			{
				\gamma_j^{j\in J}, \Pred,  
				\Post}
			{\ostate{s_i^{i\in L}} \OTWeakarrow {\alpha}
				\ostate{s_i'^{\, i\in L}}}
		}
	\]
		with  $J\cap\Holes(Q)\neq\emptyset$ or $\exists i\in L_Q.\,s_i\neq s'_i$, i.e. $Q$ takes part in the reduction.  
		 Then there exist $\gamma_{j_0}=[\beta_0..\beta_n]$, $\Pred\,'$,  
		$\Post\,'$,   and for all $p\in[0..n]$ there exist $\alpha_p$, $\Pred_p$, $\Post_p$ and a family $\gamma_{p j}^{j\in J_p}$ and for all $p\in[0..n+1]$ $s_{p i}^{\,i\in L_Q}$. s.t.:\\[-2ex]
		%\[
		\begin{mathpar}
		P\models{\openrule
			{
				\gamma_j^{j\in (J_p\setminus \Holes(Q)) \cup \{j_0\}}, 
				\Pred\,',  
				\Post\,'}
			{\ostate{s_i^{i\in L\setminus L_Q}} \OTWeakarrow {\alpha}
				\ostate{s_i'^{\,i\in L\setminus L_Q}}}
		}%\]
	\vspace{-2.2ex}\\\text{and for all $p\in[0..n]$~~}
		%\[
		Q\models{\openrule
			{
				\gamma_{p j}^{j\in J_p}, \Pred_p,  
				\Post_p}
			{\ostate{s_{p i}^{i\in L_Q}} \OTWeakarrow {\alpha_p}
				\ostate{s_{(p+1) i}^{\,i\in L_Q}}}
		}%\]
\\
		\text{and~~}  \bigcup_{p=0}^nJ_{p} = J\cap\Holes(Q) \qquad
 \text{and\qquad} \forall j\in J_p,\,\gamma_j= \mybigdotcup_{p=0}^n \gamma_{p j} \\
 \text{and\quad} \forall i \in L_Q.\, s_{(n+1) i} = s_i'\land s_{0 i} = s_i\\{\hspace{-3ex}}\text{and~~} \Pred = \Pred\,'
		\land \!\!\bigwedge_{p=0}^n((\alpha_p= \beta_p\land \Pred_p)\bigotimes_{i=0}^{p-1}\Post_i) 
\\
\text{and~}
\Post=\Post\,' \uplus \bigotimes_{p=0}^n
		\Post_p  
		\end{mathpar}
where $\Post_p$ only act upon variables of 
		$\Leaves(Q)$.
\end{lemma}

@@

Consequently, for  $\gamma_{j_0}=[\beta_0..\beta_n]$ such that $\beta_i=\alpha'_i$ we have $n$ weak open transitions  $\symb{WOT}_k$ such that $k\in [0..n]$.
Therefore, we can conclude that we have for all $k\in [0..n]$:  
\begin{mathpar}
Q\models{\openrule
			{
			\gamma_{kj}^{j\in J_k}, {\Pred_k}, {\Post_k}}
			{\ostate{s_{pl}^{l\in L_Q}} \OTWeakarrow {\alpha_{k}}
				\ostate{s_{(p+1)l}^{l\in L_Q}}}
		}%\]
\end{mathpar}
and $\displaystyle{\bigcup_{p=0}^nJ_{p}} = J\cap\Holes(Q)$, $\forall j\in J_p,\,\gamma_j= \displaystyle{\mybigdotcup_{p=0}^n \gamma_{p j}}$, $\forall l\in L_Q,\, s_{0l}=s_l$, $s_{(n+1)l}=s'_l$  \\
  $\Pred = \Pred\,'
		\land \displaystyle{\bigwedge_{p=0}^n}((\alpha_p= \beta_{pj_0}\land \Pred_p)\bigotimes_{i=p-1}^{0}\Post_i),\text{and~}$\\
$\displaystyle{\Post=\Post\,'\uplus\bigotimes_{p=n}^0
		\Post_p}$  \\
We have:\\
$\Pred=\displaystyle{\bigwedge_{i=0}^{m_1}\Pred_{1i}\subst{\bigotimes_{j=i-1}^{0}\!\!\Post_{1j}}\land\Pred_2 \subst{\bigotimes_{j=m_1}^{0}\!\!\Post_{1j}}\land 
\bigwedge_{i=0}^{m_2}\Pred_{3i}\subst{\bigotimes_{j=i-1}^{0}\!\!\Post_{3j}\shortotimes\Post_2\shortotimes\bigotimes_{j=m_1}^{0}\!\!\Post_{1j}}}$\\
By replacing $\Pred_{1i}$,  $\Pred_2$, and $\Pred_{3i}$ by their equivalent formula  we obtain:\\
$\Pred =\displaystyle{\bigwedge_{i=0}^{m_1}({\Pred'_{1i}}
		\land {\Pred''_{1i}}\land \alpha_{1i}=\beta_{1ij_0}) \bigotimes_{j=i-1}^{0}\subst{\Post_{1j}}\land}\\
~~\qquad\qquad\qquad({\Pred'_2}\land {\Pred''_2}\land \alpha_2=\beta_{2j_0})	\bigotimes_{i=m_1}^{0}\subst{\Post_{1i}}\land\\
~~\qquad\qquad\bigwedge_{i=0}^{m_2} ({\Pred'_{3i}}\land {\Pred''_{3i}}\land \alpha_{3i}=\beta_{3ij_0})
\subst{\bigotimes_{j=i-1}^{0}\!\!\Post_{3j}\shortotimes\Post_2\shortotimes\bigotimes_{j=m_1}^{0}\!\!\Post_{1j}}$\\
By distribution of the substitution operation over the logical operator we can deduce:\\
$\Pred= \displaystyle{\big(\bigwedge_{i=0}^{m_1}
{\Pred'_{1i}}\bigotimes_{j=i-1}^{0}\subst{{\Post'_{1j}}}\big)
\land{\Pred'_2}\bigotimes_{i=m_1}^{0}\subst{{\Post'_{1i}}}\land} \\
~~\qquad\qquad\quad\big(\bigwedge_{i=0}^{m_2}{\Pred'_{3i}}\bigotimes_{j=i-1}^{0}\subst{{\Post'_{3j}}}\subst{{\Post'_2}}\bigotimes_{j=m_1}^{0}\subst{{\Post'_{1j}}}\big) \land\\	
~\qquad\qquad\quad\bigwedge_{i=0}^{m_1}({\Pred''_{1i}}\land \alpha_{1i}=\beta_{1ij_0}) \bigotimes_{j=i-1}^{0}\subst{\Post_{1j}}\land({\Pred''_2}\land \alpha_2=\beta_{2j_0})	\bigotimes_{i=m_1}^{0}\subst{\Post_{1i}}\land\\
~\qquad\qquad\quad\bigwedge_{i=0}^{m_2} ({\Pred''_{3i}}\land \alpha_{3i}=\beta_{3ij_0})\bigotimes_{j=i-1}^{0}\subst{\Post_{3j}}\subst{\Post_2}  \bigotimes_{i=m_1}^{0}\subst{\Post_{1i}}
$\\		
Therefore, we have:\\
$\Pred= \Pred\,' \wedge \displaystyle{\bigwedge_{p=0}^{n}}((\alpha_p= \beta_{pj_0}\land \Pred_p)\bigotimes_{i=p-1}^{0}\Post_i)$\\ where $n=m_1+m_2+1$	\\
Concerning the postcondition we have:\\
$\Post=\displaystyle{\bigotimes_{j=m_2}^{0}\!\!\Post_{3j}\shortotimes\Post_2\shortotimes\bigotimes_{j=m_1}^{0}\!\!\Post_{1j}}$\\
By replacing ${\Post_{1i}}$,  $\Post_2$, and ${\Post_{3i}}$ and because the ${\Post'_i}$ and ${\Post''_i}$ are independent (the ${\Post'_i}$ concern the leaves of $P$ while the ${\Post''_i}$ concern the leaves of $Q$), we obtain:\\
$\Post= \displaystyle{\bigotimes_{i=m_2}^{0}\subst{{\Post'_{3i}}\uplus 
		{\Post''_{3i}}} \subst{{\Post'_2}\uplus 
		{\Post''_2}} \bigotimes_{j=m_1}^{0}\subst{ {\Post'_{1i}}\uplus 
		{\Post''_{1i}}}}\\
~~\qquad= \bigotimes_{i=m_2}^{0}{\Post'_{3i}}\subst{{\Post'_2}}\bigotimes_{j=m_1}^{0}\subst{{\Post'_{1j}}} \uplus \bigotimes_{i=m_2}^{0}{\Post''_{3i}}\subst{{\Post''_2}}\bigotimes_{j=m_1}^{0}\subst{{\Post''_{1j}}} \\
~~\qquad= \Post' \uplus \bigotimes_{j=n}^{0} \Post_j $\\
 where $n=m_1+m_2+1$	
\end{proof}
